---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(GGally)
library(rJava)
library(leaps)
library(glmulti)
```

```{r}
orange_juice <- read_csv(here::here("data/orange_juice.csv")) %>% 
  janitor::clean_names()

glimpse(orange_juice)
summary(orange_juice)
orange_juice
```

```{r}
# verify if store7 and store are redundant variables
orange_juice %>% 
  filter(store_id != store & store_id != 7)

# price_diff = sale_price_mm - sale_price_ch --> redundant
# sale_price = price - disc --> redundant
# list_price_diff = price_mm - price_ch --> redundant

# drop redundant variables and mutate categorical variables into factors
oj <- orange_juice %>%
  select(-c(weekof_purchase, store7, store, price_diff, list_price_diff, disc_ch, disc_mm)) %>% # there is a store_id already
  rename(purchase_mm = purchase) %>% 
  mutate(purchase_mm = as.factor(if_else(purchase_mm == "MM", TRUE, FALSE)),
         special_ch = as.factor(if_else(special_ch == 1, TRUE, FALSE)),
         special_mm = as.factor(if_else(special_mm == 1, TRUE, FALSE)),
         store_id = as.factor(store_id)
   )

oj
```

```{r}
alias(purchase_mm ~ ., data = oj)
```

### Automated Method  

Run an exhaustive search over all `main effects only` logistic regression models using BIC as the quality metric  


```{r}
glmulti_search_all_mains <- glmulti(
  purchase_mm ~ ., 
  data = oj,
  level = 1,               # No interactions considered, main effects only
  method = "h",            # Exhaustive approach
  crit = "bic",            # BIC as criteria
  confsetsize = 10,        # Keep 10 best models
  plotty = F, 
  report = T,              # No plots, but provide interim reports
  fitfunction = "glm",     # glm function
  family = binomial(link = "logit")) # binomial family for logistic regression

summary(glmulti_search_all_mains)
```

Best model returned : purchase_mm ~ 1 + loyal_ch + sale_price_mm + sale_price_ch  

You would like to add on a single pair interaction considering only main effects in the model. Which single pair addition leads to lowest BIC?  

```{r}
glmulti_search_previous_mains_one_pair <- glmulti(
  purchase_mm ~ 1 + loyal_ch + sale_price_mm + sale_price_ch, 
  data = oj,
  level = 2,               # Interactions considered
  method = "h",            # Exhaustive approach
  crit = "bic",            # BIC as criteria
  confsetsize = 10,        # Keep 10 best models
  marginality = TRUE,      # consider pairs only if both main effects in model
  minsize = 6,             # minsize, maxsize and marginality here force 
  maxsize = 6,             # inclusion of a single pair beyond the five main effects
  plotty = F, 
  report = T,              # No plots, but provide interim reports
  fitfunction = "glm",     # glm function
  family = binomial(link = "logit")) # binomial family for logistic regression

summary(glmulti_search_previous_mains_one_pair)
```

Best model with interactions returned : purchase_mm ~ 1 + loyal_ch + sale_price_mm + sale_price_ch + sale_price_mm:loyal_ch + sale_price_ch:loyal_ch + sale_price_ch:sale_price_mm  

Alternatively: In cases where an exhaustive search isn’t possible because there are too many possible models to search through, you could try a search using a genetic algorithm. Here, run a genetic algorithm search over all main effects plus pair models, using lowest AIC as the quality criterion.   

```{r}
# WARNING - it takes 5 minutes tu run
glmulti_ga_search_with_pairs_aic <- glmulti(
  purchase_mm ~ .,
  data = oj,
  level = 2,               # Interactions considered
  method = "g",            # Genetic algorithm approach
  crit = "aic",            # AIC as criteria
  confsetsize = 10,        # Keep 10 best models
  marginality = TRUE,      # consider pairs only if both main effects in model
  plotty = F, 
  report = T,              # No plots, but provide interim reports
  fitfunction = "glm",     # glm function
  family = binomial(link = "logit")) # binomial family for logistic regression

summary(glmulti_ga_search_with_pairs_aic)
```

After 300 generations:
Best model: purchase_mm~1+store_id+special_ch+special_mm+price_ch+price_mm+loyal_ch+sale_price_mm+sale_price_ch+pct_disc_mm+special_mm:special_ch+price_mm:price_ch+loyal_ch:price_mm+sale_price_mm:price_ch+sale_price_mm:loyal_ch+sale_price_ch:price_mm+sale_price_ch:loyal_ch+sale_price_ch:sale_price_mm+pct_disc_mm:price_ch+pct_disc_mm:price_mm+pct_disc_mm:loyal_ch+pct_disc_mm:sale_price_ch+store_id:sale_price_mm+store_id:pct_disc_mm+special_ch:sale_price_ch  

Crit= 821.572135467486
Mean crit= 822.441059302472
Improvements in best and average IC have bebingo en below the specified goals.
Algorithm is declared to have converged.
Completed.  


---  

### Manual method  

```{r}
oj_num <- oj %>% 
  select(purchase_mm, price_ch, price_mm, loyal_ch, sale_price_mm, sale_price_ch, pct_disc_mm, pct_disc_ch)

oj_fact <- oj %>% 
  select(purchase_mm, store_id, special_ch, special_mm)
```


```{r message=FALSE}
ggpairs(oj_num %>% 
          select(purchase_mm, everything()))
```

Percentage discounts and Customer brand loyalty for CH seem to be relevant.  

```{r message=FALSE}
ggpairs(oj_fact %>% 
          select(purchase_mm, everything()))
```

Special on mm/ch seem relevant  
