---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidytext)
library(tidyverse)
library(janeaustenr)
library(wesanderson)
```


### Question 1  
Find the most common words in both Pride & Prejudice and Sense & Sensibility.  


```{r}
# create a tibble of Pride and Prejudice book and unnest words
pride_book <- tibble(
  id = 1:length(prideprejudice),
  text = prideprejudice
) %>% 
  unnest_tokens(word, text)

pride_book
```

```{r}
# find the most common words in P&P
pride_book %>%
  count(word, sort = TRUE)
```

```{r}
# create a tibble of Sense and Sensibility book and unnest words
sense_book <- tibble(
  id = 1:length(sensesensibility),
  text = sensesensibility
) %>% 
  unnest_tokens(word, text)

sense_book
```

```{r}
# find the most common words in S&S
sense_book %>% 
  count(word, sort = TRUE)
```

### Question 2  
Find the most common words in both Pride & Prejudice and Sense & Sensibility, not including stop words.  

```{r}
# exclude stopwords and find the most common words in P&P
pride_book %>%
  anti_join(stop_words) %>%
  count(word, sort = TRUE)
```

```{r}
# exclude stopwords and find the most common words in S&S
sense_book %>% 
  anti_join(stop_words) %>% 
  count(word, sort = TRUE)
```

### Question 3  
Find the most common sentiment words in both Pride & Prejudice and Sense & Sensibility.  


```{r}
# save the P&P text with no stopwords
pp <- pride_book %>%
  anti_join(stop_words) 
```

```{r}
# find sentiment words in P&P using using the loughran lexicon
pp_sentiments <- pp %>%
  inner_join(get_sentiments("loughran"))

pp_sentiments
```

```{r}
# find the most common sentiment words in P&P
most_common_sentiments_pp <- pp_sentiments %>%
  count(word, sort = TRUE) %>% 
  slice_head(n = 10)

most_common_sentiments_pp
```


```{r}
# save the S&S text with no stopwords
ss <- sense_book %>%
  anti_join(stop_words) 
```

```{r}
# find sentiment words in S&S using using the loughran lexicon
ss_sentiments <- ss %>%
  inner_join(get_sentiments("loughran"))

ss_sentiments
```

```{r}
# find the most common sentiment words in S&S
most_common_sentiments_ss <- ss_sentiments %>%
  count(word, sort = TRUE) %>% 
  slice_head(n = 10)

most_common_sentiments_ss
```


### Extension  
Taking your results above, can you create a plot which visualises the differences between the books?  


```{r}
# pull a vector of the above sentiment words in P&P to use as a filter
most_common_sentiments_pp_words = most_common_sentiments_pp %>% pull(word)
# plot them to visualise how many times they occur, and associated sentiments
pp_sentiments %>% 
  filter(word %in% c(most_common_sentiments_pp_words)) %>% 
  mutate(word = factor(word, levels = most_common_sentiments_pp_words)) %>%
  ggplot() +
  aes(x = word, fill = sentiment) +
  geom_bar() +
  scale_fill_manual(values = wes_palette("FantasticFox1"))
```

```{r}
# pull a vector of the above sentiment words in S&S to use as a filter
most_common_sentiments_ss_words = most_common_sentiments_ss %>% pull(word)
# plot them to visualise how many times they occur, and associated sentiments
ss_sentiments %>% 
  filter(word %in% c(most_common_sentiments_ss_words)) %>% 
  mutate(word = factor(word, levels = most_common_sentiments_ss_words)) %>%
  ggplot() +
  aes(x = word, fill = sentiment) +
  geom_bar() +
  scale_fill_manual(values = wes_palette("FantasticFox1"))
```

```{r}
# save a tibble of all most common sentiments words in P&P
# add a column with book name to use in facet_wrap() afterwards
most_common_sentiments_pp_words_all <- pp_sentiments %>% 
  filter(word %in% c(most_common_sentiments_pp_words)) %>% 
  mutate(word = factor(word, levels = most_common_sentiments_pp_words),
         book = "Pride and Prejudice")

# save a tibble of all most common sentiments words in S&S
# add a column with book name to use in facet_wrap() afterwards
most_common_sentiments_ss_words_all <- ss_sentiments %>% 
  filter(word %in% c(most_common_sentiments_ss_words)) %>% 
  mutate(word = factor(word, levels = most_common_sentiments_ss_words),
         book = "Sense and Sensibility")

# bind the two tibbles
compare_pp_ss <- rbind(most_common_sentiments_pp_words_all, most_common_sentiments_ss_words_all)
```

```{r}
# compare the most common sentiment words in the two books (number of occurrences and associated sentiments) 
compare_pp_ss %>%
  ggplot() +
  aes(x = word, fill = sentiment) +
  geom_bar() +
  facet_wrap(~ book) +
  scale_fill_manual(values = wes_palette("FantasticFox1")) +
  theme(axis.text.x = element_text(size = 8, angle = 45, vjust = 1, hjust = 1))
```
