---
title: "R Notebook"
output: html_notebook
---

```{r}
library(janitor)
library(broom)
library(tidyverse)
library(factoextra)
```

```{r}
customers <- read_csv(here::here("data/mall_customers.csv")) %>% 
  clean_names()

customers
```

### Question 1  

Do some exploratory analysis on the data. Comment on findings.

```{r}
summary(customers)

# check if we have NAs
customers %>% 
  summarise(across(everything(), .fns = ~sum(is.na(.x))))
```

```{r}
customers %>% 
  ggplot() +
  aes(x = age, y = annual_income_k) +
  geom_point()
# no clear relationship between age and annual income

customers %>% 
  ggplot() +
  aes(x = age, y = spending_score_1_100) +
  geom_point()
# some patterns between age and spending score, customers under 40 have higher spending score than those over 40

customers %>% 
  ggplot() +
  aes(x = annual_income_k, y = spending_score_1_100) +
  geom_point()
# can see 5 different types of behaviours here

customers %>% 
  ggplot() +
  aes(x = annual_income_k, y = spending_score_1_100) +
  geom_point() +
  facet_wrap(~gender)
# not much difference between male and female, we still see 5 groups
```
The relationship between annual income and spending score seems promising in terms of having different groups of behaviour.  


### Question 2  

We are interested in creating a marketing campaign to target customers based on their spending score and annual income. Perform a k-means clustering to find if there are meaningful clusters in the data to target the customers.  

```{r}
customers_sub <- customers %>% 
  select(annual_income_k, spending_score_1_100)


customers_sub %>% 
  ggplot(aes(x = annual_income_k, y = spending_score_1_100)) +
  geom_point()  
```
Can definitely identify 5 clusters or maybe more  


```{r}
# check if needs scaling
customers_sub %>%
  as_tibble() %>%
  pivot_longer(cols = c(annual_income_k, spending_score_1_100), 
               names_to = "type", 
               values_to = "value") %>% 
  group_by(type)%>%
  summarise(mean = round(mean(value)), 
            sd = sd(value))

# scaling
customers_scale <- customers_sub %>%
                mutate_if(is.numeric, scale)

# checking if scaled
customers_scale %>%
  as_tibble() %>%
  pivot_longer(cols = c(annual_income_k, spending_score_1_100), 
               names_to = "type", 
               values_to = "value") %>% 
  group_by(type)%>%
  summarise(mean = round(mean(value)), 
            sd = sd(value))
```




1- Elbow method

```{r}
# library(factoextra)
fviz_nbclust(customers_scale, 
             kmeans, 
             method = "wss", 
             nstart = 25)
```
This graph is smoother than we want it to be - there is a kind of an elbow around 3 and 5   


2 - Silhouette coefficient Method

```{r}
fviz_nbclust(customers_scale, 
             kmeans, 
             method = "silhouette", 
             nstart = 25)
```

Definitely better - we've got something at 5, but 6 and 7 are not very far    


3- Gap statistic

```{r}
fviz_nbclust(customers_scale, 
             kmeans, 
             method = "gap_stat",
             nstart = 25)
```

This gives us k = 1 (confusing different results!)  
It may mean the data is not suitable for k-means clustering  



### Question 3  

Perform k-means clustering and chose a value of k.  

```{r}
# choosing 5 as it was my initial intuition when plotting the 2 variables
clustered_customers <- kmeans(customers_scale, 
                              centers = 5, 
                              nstart = 25)

clustered_customers
```
```{r}
clustered_customers$size

clustered_customers$centers
```





### Question 4  

Visualise the clustering for your chosen value of k.  

```{r}
clusters <- augment(clustered_customers, customers)

ggplot(clusters, aes(x = annual_income_k, y = spending_score_1_100, colour = .cluster)) +
  geom_point() 
```
 

### Question 5  

Do you think the clustering seems a good fit for this data? 

Tried also with 7 clusters and as expected the data point for  x > 100 fall into 2 more groups (above and below y = 50).  

In spite of the different results using the 3 different methods, I believe the initial intuition of 5 clusters still holds.  




### Question 6  

Comment on the attributes on one or two of the clusters (maybe even give them a label if you like - like in section 4.1 of the ‘Segmentation & clustering intro’ lesson).  

For annual income > 60k we can see 2 very different types of behaviours:  
1) The wise buyer - their spending score ranges between 0 and 40 (on a scale from 0 to 100), despite their income getting fairly high    
2) The wild customer - their spending score is > 60 (out of 100) and they definitely enjoy their life  

