---
title: "Cancer Incidence across NHS Borders"
author: "Manuela Mastrogiovanni"
date: "2022-12-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, message=FALSE}
library(tidyverse)
library(janitor)
library(slider)
# HB codes
hb <- read_csv("raw_data/geography_codes_and_labels_hb2014_01042019.csv")
# all cancers for each HB for 1996-2020, all ages grouped
incidence_hb <- read_csv("raw_data/opendata_inc9620_hb.csv") %>% 
  clean_names()
# all cancers for HB = Borders, for 1996-2020, all ages grouped
borders_1996_2020 <- incidence_hb %>% 
  filter(hb == "S08000016")
```

---

### Business Requirements  

This report provides an insight of the incidence of cancer in NHS Borders, with the aim to help inform the planning for provision of cancer treatment services.  
The data was provided by Public Health Scotland ([link](https://www.opendata.nhs.scot/dataset/annual-cancer-incidence)), which is Scotland's national health improvement agency and one of the country's main national public health bodies.

```{r echo=FALSE, message=FALSE, warning=FALSE}
borders_1996_2020 %>% 
  filter(sex == "All" & cancer_site == "All cancer types") %>% 
  mutate(
    avg_incidence = slide_dbl(
      .x = incidences_all_ages, 
      .f = ~ mean(., na.rm = TRUE),
      .before = 1,
      .after = 1
    )
  ) %>% 
  ggplot() +
  geom_line(aes(x = year, y = incidences_all_ages), colour = "steelblue", alpha = 0.5) +
  geom_line(aes(x = year, y = avg_incidence), colour = "firebrick") +
  labs(title = "Incidence of New Cancer Registrations", subtitle = "NHS Borders (1996-2020)", x = "Year of diagnosis", y = "Total number of new cases") 
```

Note how after the **highest peak in 2017**, the incidence of new cancer cases have significantly decreased from 2018 onwards.



```{r echo=FALSE, message=FALSE, warning=FALSE}
borders_1996_2020 %>%
  filter(sex != "All" & cancer_site == "All cancer types") %>%
  ggplot() +
  aes(x = year, y = incidences_all_ages, fill = sex) +
  geom_col(stat = "count", position = "dodge") +
  scale_fill_manual(
    values = c(
      "Female" = "#fc8d62",
      "Male" = "#66c2a5"
      )
    ) +
  labs(title = "Incidence of New Cancer Registrations by Sex", subtitle = "NHS Borders (1996-2020)", x = "Year of diagnosis", y = "Total number of new cases", fill = "Sex") +
  theme(panel.grid = element_blank())
```

On average, **women have been more affected by cancer than men**. This trend changed, in particular between 2015 and 2018.



```{r echo=FALSE, message=FALSE, warning=FALSE}
borders_1996_2020 %>% 
  filter(cancer_site != "All cancer types" & sex == "All") %>% 
  ggplot() +
  aes(x = year, y = incidences_all_ages, colour = cancer_site) +
  geom_line() +
  facet_wrap(
    ~ cancer_site,
    scales = "free",
    labeller = as_labeller(
      function(x) {
        str_trunc(x, 15)
      }
    )
  ) +
  labs(title = "Trends by Cancer Type", subtitle = "NHS Borders (1996-2020)", x = "Timeline", y = "New cancer cases", colour = "Cancer Type") + 
  theme(legend.position = "none", 
        strip.text.x = element_text(size = 8, hjust = 0),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.grid = element_blank())
```


```{r echo=FALSE, message=FALSE, warning=FALSE}
borders_1996_2020 %>% 
  filter(cancer_site != "All cancer types" & sex == "All") %>% 
  group_by(cancer_site) %>% 
  summarise(total = sum(incidences_all_ages)) %>% 
  arrange(desc(total))
```

Cancer types that have seen a steep increase over time:  
- Lip & oral cavity, oropharyngeal, trachea, broncus, lung  
- Liver  
- Malignant and non melanoma skin cancer  
- Pancreas  
- Colon & colonrectal   