---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(modelr)
library(GGally)
library(ggfortify)
```

```{r}
avocado <- read_csv(here::here("data/avocado.csv")) %>% 
  janitor::clean_names()
```

```{r}
glimpse(avocado)
summary(avocado)
avocado %>% 
  select(-x1) %>% 
  skimr::skim()
```
All the variables seem to be right;ly skewed apart prom the average_price.  
The total_volume represents the total sale of avocado for each week: total_volume = x4046 + x4225 + x4770 + total_bags (total_bags = small_bags + large_bags + x_large_bags), therefore I expect a few of these variables to be redundant. The alias() should be able to detect which one can be dropped.  

```{r}
alias(average_price ~ ., data = avocado)
```


```{r}
avocado_extended <- avocado %>% 
  mutate(is_organic = if_else(type == "organic", TRUE, FALSE),
         month = lubridate::month(date, label = TRUE, abbr = TRUE)
  ) %>% 
  select(-c(region, type))

avocado_extended_with_mean <- avocado_extended %>% 
  mutate(week = lubridate::week(date)) %>%
  group_by(week, is_organic) %>% 
  mutate(avg_price_per_week = mean(average_price))

avocado_extended_with_mean %>% 
  filter(year == 2015) %>% 
  group_by(week) %>% 
  mutate(avg = mean(avg_price_per_week)) %>% 
  arrange(week)
```

```{r}
avocado_extended_with_mean %>% 
  ggplot(aes(x = week, y = avg_price_per_week, group = year, colour = is_organic)) +
  geom_line() +
  facet_wrap(~ is_organic)
```

```{r}
alias(average_price ~ ., data = avocado_extended_with_mean)
```

```{r}
avocado_test <- avocado_extended %>% 
  select(-c(x1, date))
```

```{r message=FALSE}
ggpairs(avocado_test)
```
We appreciate 2 different behaviours in average_price when looking at organic and conventional avocados.  

```{r}
avocado_test %>%
  ggplot(aes(x = is_organic, y = average_price)) + 
  geom_boxplot()
```

The distributions of FALSE and TRUE are different (avg_price looks higher for organic avocados).  There is evidence to investigate whether or not there is an association between the two variables.  

```{r}
avocado_test %>%
  summarise(cor = cor(is_organic, average_price))
```

```{r}
model1 <- lm(average_price ~ is_organic, data = avocado_test)
summary(model1)
autoplot(model1)
```

```{r}
avocado_test %>%
  summarise(cor = cor(total_volume, average_price))
```



```{r}
model2 <- lm(average_price ~ total_volume, data = avocado_test)
summary(model2)
autoplot(model2)
```

