---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
books <- read_csv("data/books.csv")
```

# Q0: Investigate the dataset

```{r}
spec(books)

ncol(books)

nrow(books)
      
dim(books) 

names(books)

head(books)

tail(books)

view(books)

books
      
```


# Q1 : Show all languages ordered by book count 

```{r}
# select the columns to keep
book_study <- select(books, bookID, title, authors, average_rating, language_code, publication_date)


# how many languages there are?
book_study %>%
  group_by(language_code) %>% 
  summarise(book_count = n()) %>% 
  arrange(desc(book_count))
```


# Q2 : Show all english language authors in the dataset, sorted by descending book count

```{r}
book_study %>%
  filter(language_code == "eng") %>% 
  group_by(authors) %>% 
  summarise(book_count = n()) %>% 
  arrange(desc(book_count))
```


# Q3: Find the 3 authors (eng) with the highest average rating

```{r}
book_study %>%
  filter(language_code == "eng") %>% 
  group_by(authors) %>% 
  summarise(average_rating = mean(average_rating, na.rm = TRUE)) %>% 
  slice_max(average_rating, n = 3, with_ties = FALSE)

# with_ties = TRUE I get all the authors with average_rating = 5 (i.e. 16)
```


# Q4: Separate and count the english books from all other languages

```{r}
book_study %>% 
  select(language_code) %>% 
  mutate(language_code = recode(language_code,"en-US" = "eng", "en-GB" = "eng", "en-CA" = "eng", "eng" = "eng", .default = "other")) %>% 
  group_by(language_code) %>% 
  summarise(book_count = n())
```




# Q5: Separate and count the english books from all other languages using case_when

```{r}
book_study %>% 
  select(language_code) %>% 
  mutate(language_code = case_when(
    substr(language_code, 1, 2) == "en" ~ "eng",
    TRUE ~ "other"
    )
  ) %>% 
  group_by(language_code) %>% 
  summarise(book_count = n())

# tried using starts_with, could not use in this case, found substr()
```


```{r}
# when noticed the book_count returned was different from the one in Q4 (because of the enm language included) I added the string language_code == "eng" ~ "eng" 

book_study %>% 
  select(language_code) %>% 
  mutate(language_code = case_when(
    substr(language_code, 1, 3) == "en-" ~ "eng",
    language_code == "eng" ~ "eng",
    TRUE ~ "other"
    )
  ) %>% 
  group_by(language_code) %>% 
  summarise(book_count = n())

```





# Q Optional: Order the books by publication date, group them by decades, assess the average rating for each decade, assess the decades with the highest and lowest average rating . Find the decades with most and least books published


```{r}
# as.Date()
# not clear how dates work 
```

