---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(GGally)

kc_house <- read_csv(here::here("data/kc_house_data.csv"))
kc_house
summary(kc_house)
skimr::skim(kc_house)
```

### Question 1  
* You might like to think about removing some or all of `date`, `id`, `sqft_living15`, `sqft_lot15` and `zipcode` (`lat` and `long` provide a better measure of location in any event).  
* Have a think about how to treat `waterfront`. Should we convert its type?  
* We converted `yr_renovated` into a `renovated` logical variable, indicating whether the property had ever been renovated. You may wish to do the same.  
* Have a think about how to treat `view`, `condition` and `grade`. Are they interval or categorical ordinal data types?  

```{r}
kc_house_trim <- kc_house %>% 
  select(-c(id, date, zipcode, sqft_living15, sqft_lot15)) %>% 
  rename(renovated = yr_renovated) %>% 
  mutate(waterfront = if_else(waterfront == 1, TRUE, FALSE),
         renovated = if_else(renovated != 0, TRUE, FALSE),
         view = case_when(view == 0 ~ "NONE",
                          view == 1 ~ "POOR",
                          view == 2 ~ "AVERAGE",
                          view == 3 ~ "GOOD",
                          view == 4 ~ "VERY GOOD"),
         condition = case_when(condition == 1 ~ "VERY POOR",
                               condition == 2 ~ "POOR",
                               condition == 3 ~ "AVERAGE",
                               condition == 4 ~ "GOOD",
                               condition == 5 ~ "VERY GOOD"),
         grade = case_when(grade >= 11 ~ "VERY HIGH",
                           grade >= 9 ~ "HIGH",
                           grade >= 6 ~ "AVERAGE",
                           grade >= 4 ~ "POOR",
                           TRUE ~ "VERY POOR")
         )

kc_house_trim
```

Unsure whether to use the dummy_cols() function to treat the categorical ordinal data types (view, condition, grade).  

### Question 2  
* Check for aliased variables using the alias() function (this takes in a formula object and a data set).  
[Hint - formula price ~ . says ‘price varying with all predictors’, this is a suitable input to alias()].  
* Remove variables that lead to an alias. Check the ‘Elements of multiple regression’ lesson for a dropdown containing further information on finding aliased variables in a dataset.  

```{r}
alias(lm(price ~ ., data = kc_house_trim))
```
The sqft_basement can be deduced by calculating sqft_living - sqft_above, therefore it can be dropped.  

```{r}
kc_house_trim <- kc_house_trim %>% 
  select(-sqft_basement)

kc_house_trim
```

### Question 3  
Systematically build a regression model containing up to four main effects (remember, a main effect is just a single predictor with coefficient), testing the regression diagnostics as you go and the same in subsequent rounds of predictor selection with the resid column.   
[Hint - Splitting datasets into numeric and non-numeric columns might help ggpairs() run in manageable time, although you will need to add either a price or resid column to the non-numeric dataframe in order to see its correlations with the non-numeric predictors].  

Remember, if you are not sure whether including a categorical predictor is statistically justified, run an anova() test passing in the models with- and without the categorical predictor and check the p-value of the test. 

```{r message=FALSE}
houses_tidy_numeric <- kc_house_trim %>%
  select_if(is.numeric)

houses_tidy_nonnumeric <- kc_house_trim %>%
  select_if(function(x) !is.numeric(x))

houses_tidy_nonnumeric$price <- kc_house_trim$price

ggpairs(houses_tidy_numeric)

houses_tidy_nonnumeric %>% 
  select(price, everything()) %>% 
  ggpairs(aes(colour = condition, alpha = 0.5))
```

*Numeric*  
The house price is strongly positively correlated to:  
1 - sqft_living  
2 - sqft_above  
3 - bathrooms  
4 - bedrooms  

*Non numeric*
