---
title: "R Notebook"
output: html_notebook
---

```{r}
library(e1071)
library(infer)
library(janitor)
library(skimr)
library(tidyverse)
```




## Task 1
Load the data, clean_names(), and familiarise yourself with it.

```{r}
ames <- read_csv(here::here("data/ames.csv")) %>% 
  clean_names()

glimpse(ames)
head(ames)
```



## Task 2
Investigate the distribution of lot_area. Is the distribution roughly normal? If not, what problems do you find?

```{r}
ames %>% 
  ggplot(aes(x = lot_area)) +
  geom_histogram(col = "white", 
                 bins = 50,
                 fill = "steelblue",
                 alpha = 0.7)
```

```{r}
# assess skewness
ames %>% 
  summarise(skew = skewness(lot_area, type = 1))

# five number summary
ames %>% 
  select(lot_area) %>% 
  skim()

# assess outliers with a box plot
ames %>% 
  ggplot(aes(x = lot_area)) +
  geom_boxplot()
```

The distribution is unimodal, non-normal, highly right-skewed.
The range is strongly affected by outliers.



# Task 3
Compute and visualise a bootstrap sampling distribution for the mean(lot_area) of the sold houses.

```{r}
# bootstrap distribution workflow and visualisation using the "infer" package  
ames_infer_resample <- ames %>% 
  specify(response = lot_area) %>% 
  generate(reps = 5000, type = "bootstrap") %>% 
  calculate(stat = "mean")

ames_infer_resample %>%
  visualise(bins = 30)
```


# Task 4
Use your bootstrap distribution to calculate a 95% CI for mean(lot_area), and visualise it on the distribution

```{r}
ames_infer_ci_95 <- ames_infer_resample %>%
  get_confidence_interval(level = 0.95, type = "percentile")

ames_infer_ci_95

ames_infer_resample %>%
  visualise(bins = 30) +
  shade_confidence_interval(endpoints = ames_infer_ci_95)
```



## Task 5
You would like to know the mean(lot_area) of the sold houses with higher confidence. Calculate the 99% CI for this variable (you can re-use your bootstrap distribution from above). Is it narrower or broader than the 95% CI? Does that make sense?

```{r}
ames_ci_99 <- bootstrap_resample_2930 %>%
  summarise(
    mean = mean(mean_lot_area),
    lower_bound = quantile(mean_lot_area, probs = 0.005),
    upper_bound = quantile(mean_lot_area, probs = 0.995)
  )

ames_ci_99
```

The 99% CI for the mean(lot_area) is wider than the 95% CI, as expected (9817.23-10544.93 as opposed to 9876.433-10456.16).



## Task 6
Calculate the point estimate of the mean(lot_area).

```{r}
mean_infer <- ames_infer_resample %>%
  summarise(mean = mean(stat))

mean_infer
```

```{r}
# compared to the mean of the original data set, they are very close
ames %>%
  summarise(mean(lot_area))
```



## Extension
Calculate a point estimate and 95% CI for the proportion of houses in the data built before 1920. 
Does the number of reps you use matter? [Investigate reps from 200 up to 50000, memory of your laptop permitting].

```{r}
ames_before_1920 <- ames %>% 
  filter(year_built < 1920)

# investigate the distribution of sale_price
ames_before_1920 %>% 
  ggplot(aes(x = sale_price)) +
  geom_histogram(col = "white", 
                 bins = 50,
                 fill = "steelblue",
                 alpha = 0.7)

options(scipen = 5)
```


```{r}
ames_before_1920_infer_resample <- ames_before_1920 %>% 
  specify(response = sale_price) %>% 
  generate(reps = 200, type = "bootstrap") %>% 
  calculate(stat = "mean")

ames_before_1920_infer_resample %>%
  visualise(bins = 30)

ames_before_1920_infer_ci_95 <- ames_before_1920_infer_resample %>%
  get_confidence_interval(level = 0.95, type = "percentile")

ames_before_1920_infer_ci_95

ames_before_1920_infer_resample %>%
  visualise(bins = 30) +
  shade_confidence_interval(endpoints = ames_before_1920_infer_ci_95)
```
