---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
```


```{r}
movies_data <- read_csv(
  "data/movies.csv"
)

movies_data
```
## Q1: write down any general thoughts you get looking at the data

# do title and original_title have the same values all over?
# how are they sorted?
# columns are too wide
# "0" entered in budget
# overview
# "0" entered in revenue


## Q2: Select the three columns that contain the movies’ titles, their runtime as well as their budget and save them into a new variable called movies_selected.


```{r}
movies_selected <- select(movies_data, title, runtime, budget)

movies_selected
```

## Q3: Count the number of missing values in each column.


```{r}
movies_selected %>%
  summarise(count = sum(is.na(title)))
```

```{r}
movies_selected %>%
  summarise(count = sum(is.na(runtime)))
```

```{r}
movies_selected %>%
  summarise(count = sum(is.na(budget)))
```

## Q4: use the function na_if to convert runtimes of length 0 into NAs. Save this into a new variable called movies_runtime. Then count how many missing values there are now in the column runtime.

```{r}
movies_runtime <- movies_selected %>%
  mutate(runtime = na_if(runtime, 0))

movies_runtime
```


```{r}
movies_runtime %>%
  summarise(count = sum(is.na(runtime)))
```



## Q5: Impute the missing values in runtime with the median runtime using coalesce and save this into a new variable called movies_imputed. There should now be no missing values in the column runtime - check this to make sure.

```{r}
med_runtime <- movies_runtime %>%
  arrange(desc(runtime)) %>%
  pull(runtime) %>%
  median(runtime, na.rm = TRUE)

med_runtime

# movies_imputed <- mutate(coalesce(runtime, median())

                         
```



## Q6: use the slice_min and slice_max functions on movies_imputed to get the movies with the 10 shortest and 10 longest runtimes.


## Q7: Using the if_else function, impute anything with a budget below $100 with the median budget. Overwrite your movies_imputed so it contains the new budget values.