---
title: "R Notebook"
output: html_notebook
---


```{r}
library(CodeClanData)
library(janitor)
library(tidyverse)
```

```{r}
qb_revenue_breakdown <- clean_names(qb_revenue_breakdown)
qb_monthly_sales <- clean_names(qb_monthly_sales)
qb_competitors <- clean_names(qb_competitors)
qb_device_data <- clean_names(qb_device_data)
```


## Part 1
### Question 1
Take the data in the dataset qb_revenue_breakdown and make a stacked bar chart showing the sources of revenue across the two years in the dataset.


```{r}
 ggplot(qb_revenue_breakdown) +
  geom_col(aes(x = year, y = revenue, fill = product), position = "dodge")
```



### Question 2
Make a line chart showing monthly sales of the “flagship product” - the ‘KwikBit Unit 1’ - over the last year. This data is inside qb_monthly_sales, but you will need to use filter() or subset() (base R).


```{r}
qb_monthly_sales %>% 
  filter(date >= "2019-01-01" & date <= "2019-12-31", cashflow == "Kwikbit Unit 1 Sales") %>% 
  arrange(date) %>% 
  ggplot() +
  geom_line(aes(x = date, y = amount, colour = cashflow)) +
  geom_point(aes(x = date, y = amount))
  
```




### Question 3
Make a line chart showing monthly revenue and costs over the last year. This data is also in qb_monthly_sales.

```{r}
qb_monthly_sales %>% 
  filter(date >= "2019-01-01" & date <= "2019-12-31", cashflow !="Kwikbit Unit 1 Sales") %>% 
  arrange(date) %>% 
  ggplot() +
  geom_line(aes(x = date, y = amount, group = cashflow, colour = cashflow)) +
  geom_point(aes(x = date, y = amount, group = cashflow, colour = cashflow))
```



### Question 4
Show annual sales of personal fitness trackers over the last 5 years broken down by company as a ribbon plot (use geom_area). This data is in qb_competitors.


```{r}
qb_competitors %>% 
  ggplot() +
  geom_area(
    aes(x = year, y = revenue, group = company, colour = company), 
    fill = "grey70", alpha = .5, orientation = "x"
  ) +
  scale_y_continuous(labels = scales::comma) 
  
```




### Question 5
Now show the sales from the four competitors as a line graph. Include an extra layer that shows the data points used to make the lines.

```{r}
qb_competitors %>% 
  ggplot() +
  geom_line(aes(x = year, y = revenue, group = company, colour = company)) +
  geom_point(aes(x = year, y = revenue, group = company, colour = company)) +
  scale_y_continuous(labels = scales::comma) 
   
```




### Question 6
Now the company wants to compare the number of steps that their device counts vs. their competitors. Make a line graph of the number of steps throughout time, and use faceting to compare between companies and people. The data for this is in qb_device_data.


```{r}
qb_device_data %>% 
  mutate(tot_mins = hours * 60 + mins) %>% 
  ggplot() +
  aes(x = tot_mins, y = counts, group = device, colour = device) +
  geom_line() +
  facet_grid(device ~ id)
```



## Part 2
### Question 1
Take the plots that you produced in part one and now polish them by:

Adding appropriate labels
Changing the scales and coordinates when appropriate.
Applying a unified theme, which is described below:
Graphs should have white backgrounds, and use colour sparingly.
There should be faint grid lines.
Font sizes should be ~12pt although titles should be slightly larger and axis labels can be slightly smaller.
All plots should use colours from the following company colour scheme.