---
title: "R Notebook"
output: html_notebook
---

```{r}
library(gtools)
library(janitor)
library(tidyverse)
source("prob.R")
```



```{r}
titanic <- read_csv("data/titanic.csv")
```


### Question
Calculate the conditional probabilities prob(survived∣crew) and prob(survived∣passenger) to settle this dispute.

Have a look at the distinct values of the class variable, this will tell you whether each individual was a passenger (1st, 2nd or 3rd class) or a member of the crew (treat restaurant staff as crew).

Hints :
- You may want to create a crew column containing values TRUE or FALSE to help with this.
- For prob(survived∣crew) think ‘what proportion of the crew survived?’, and similar for prob(survived∣passenger).


```{r}
titanic %>% 
  distinct(class)
```

```{r}
titanic_survivals <- titanic %>%
  mutate(crew = case_when(
    class %in% c("3rd", "2nd", "1st") ~ FALSE,
    TRUE ~ TRUE
  )) 
```

```{r}
titanic_table <- titanic_survivals %>%
  tabyl(survived, crew)

titanic_table
```

```{r}
titanic_table <- titanic_table %>%
  adorn_totals(where = c("row", "col"))

titanic_table
```

```{r}
titanic_table <- titanic_table %>%
  adorn_percentages(denominator = "all") %>%
  adorn_rounding(digits = 4) %>% 
  adorn_title()

titanic_table
```

```{r}
prob_surv_given_crew <- titanic_survivals %>% 
  filter(crew == "TRUE") %>% 
  summarise(prob = sum(survived == "yes") / n())

prob_surv_given_crew
```

p(survived | crew) = 0.096 /0.403 = 0.2370679


```{r}
prob_surv_given_passenger <- titanic_survivals %>% 
  filter(crew == "FALSE") %>% 
  summarise(prob = sum(survived == "yes") / n())

prob_surv_given_passenger
```

p(survived | passenger) = 0.22655 /0.59674 = 0.3796507



### Question 1
If I roll a six-sided die once, what is the probability of getting a number less than 5?


```{r}
s_die <- rolldie(1)

s_under_five <- s_die %>% 
  filter(X1 < 5)

prob_under_five <- nrow(s_under_five) / nrow(s_die)

prob_under_five
```



### Question 2
How many outcomes are there in the sample space of an experiment in which we toss three dice?


```{r}
s_three_dice <- rolldie(3)

outcome_three_dice <- s_three_dice %>% 
  nrow()

outcome_three_dice
```



### Question 3
In an experiment where we roll four dice:

- how many outcomes are in the event where the total of the four dice equals 5, 6 or 7)?

Hint : For part i, rowSums(.) may help you setup up the necessary random variable total (or you can use pivot_longer())


```{r}
s_four_dice <- rolldie(4)

outcome_four_dice_5_6_7 <- s_four_dice %>% 
  filter((X1 + X2 + X3 + X4) %in% c(5:7)) %>% 
  nrow()

outcome_four_dice_5_6_7
```


- what is the probability of this event?

```{r}
prob_four_dice_5_6_7 <- outcome_four_dice_5_6_7 / nrow(s_four_dice)

prob_four_dice_5_6_7
```


- what assumptions do we apply in answering part ii?

We assume that the order does not matter.



### Question 4
Five chairs are placed in a row in a room, along with five people of differing height. The people sit down randomly on the chairs. What is the probability that they sit in order of decreasing height, left to right?

Hint : This is an arrangement where order matters. Of all the possible arrangements, how many are correct?


```{r}
five_people_height <- c(150, 160, 170, 180, 190)

five_people_combination <- as_tibble(
  permutations(
    n = 5, 
    r = 5, 
    v = five_people_height, 
    repeats.allowed = FALSE)
  )

five_people_combination <- five_people_combination %>%
  rename(
    chair_1 = V1,
    chair_2 = V2,
    chair_3 = V3,
    chair_4 = V4,
    chair_5 = V5
  )

five_people_combination

# n of permutations
nrow(five_people_combination)
```
```{r}
# get the event space
ascend_order <- five_people_combination %>%
  filter(chair_1 == 150, 
         chair_2 == 160,
         chair_3 == 170,
         chair_4 == 180,
         chair_5 == 190)

ascend_order

# calculate the probability
prob <- nrow(ascend_order) / nrow(five_people_combination)

prob
```



### Question 5
I roll six dice:

- what is the probability I see the numbers 6, 5, 4, 3, 2 and 1 in order on the dice, left to right?
- what is the probability I see one each of the numbers 6, 5, 4, 3, 2 and 1 on the dice regardless of order?

Hint : Have a look at the isin() function in the prob.R file. For example, to filter all rows containing 6, 1, 2 in that order, you could use isin(., c(6, 1, 2), ordered = TRUE).



### Question 6
If I draw a single card from a 52-card deck, and it is black, what is the probability that its suit is spades?

Hints :
- Investigate the cards() function in the prob.R file, if you’d like to try this in R (make a deck without jokers).
- You could think about adding your own black_card random variable to the sample space, which would be TRUE if the suit is ‘Spade’ or ‘Club’, and FALSE otherwise.