---
title: "R Notebook"
output: html_notebook
---


### Question 1

Load in the libraries you need, and look at the backpack data. Take note of variable names and dimensions.

```{r}
library(CodeClanData)
library(tidyverse)
```

```{r}
glimpse(backpack)
```
```{r}
head(backpack)
```


### Question 2

What is the distribution of backpack weights that students are carrying? Choose an appropriate graph to show the distribution.

```{r}
ggplot(backpack) +
  aes(x = backpack_weight) +
  geom_histogram(bins = 17, col = "white") +
  scale_x_continuous(breaks = 1:35) + 
  scale_y_continuous(breaks = c(5, 10, 15, 20)) +
  labs(
    x = "Backpack Weight (pounds)",
    y = "Count",
    title = "Backpack Weight Distribution",
    subtitle = "Data from students at California Polytechnic State University"
  )
```

The histogram shows the highest frequency of backpack weight is between 9 and 11 pounds, whereas there is a gap between 22 and 23 pounds and between 26 and 34 pounds. 


### Question 3

Are there differences in backpack weight between sex?

```{r}
ggplot(backpack) +
  aes(x = backpack_weight, fill = sex) +
  geom_histogram(bins = 17, col = "white") +
  scale_x_continuous(breaks = seq(from = 5, to = 35, by = 5)) + 
  scale_y_continuous(breaks = c(5, 10, 15, 20)) +
  facet_wrap(~ sex) +
  labs(
    x = "Backpack Weight (pounds)",
    y = "Count",
    title = "Backpack Weight Distribution",
    subtitle = "Data from students at California Polytechnic State University"
  )
```

The distribution shows that women tend to carry lighter backpacks (the majority of them between 2 and 15 lb), whereas men carry heavier backpacks (up to 35lb).


### Question 4

What is the relationship between body weight and backpack weight?

```{r}
backpack %>% 
  ggplot() +
  geom_point(aes(x = body_weight, y = backpack_weight)) +
  geom_smooth(method = lm, aes(x = body_weight, y = backpack_weight)) +
  labs(
    x = "Body Weight (pounds)",
    y = "Backpack Weight (pounds)",
    title = "Relationship between Body Weight and Backpack Weight",
    subtitle = "Data from students at California Polytechnic State University"
  )
```

The chart shows that there is a slight increase in backpack weight when the student's body weight increases too. 
However, there is no direct relationship (students with higher weight still carry light backpacks).


### Question 5

Is this relationship different between men and women? Choose an appropriate plot and interpret your findings.


```{r}
backpack %>% 
  ggplot() +
  geom_point(aes(x = body_weight, y = backpack_weight, colour = sex)) +
  geom_smooth(method = lm, aes(x = body_weight, y = backpack_weight, colour = sex)) +
  facet_wrap(~ sex) +
  labs(
    x = "Body Weight (pounds)",
    y = "Backpack Weight (pounds)",
    title = "Relationship between Body Weight and Backpack Weight",
    subtitle = "Data from students at California Polytechnic State University"
  )
```

The ratio between the student's body weight and their backpack weight is in a more direct proportion among women when compared to men.


### Question 6

What proportion of students report back pain?

```{r}
backpack %>% 
  mutate(back_problems = ifelse(back_problems == 1, "Yes", "No")) %>% 
  ggplot(aes(x = back_problems, fill = back_problems)) +
  geom_bar(aes(y = (..count..)/sum(..count..))) +
  # geom_text(aes(y = ?, label = ?)) +
  scale_y_continuous(labels = scales::percent) +
  labs(
    x = "Back Problems",
    y = "",
    title = "Reported Back Pain",
    subtitle = "Data from students at California Polytechnic State University",
    fill = "Back Pain"
  )
```

78% of the students reported back pain against 32% who did not.


### Question 7

Is this proportion different between men and women?


```{r}
backpack %>% 
  mutate(back_problems = ifelse(back_problems == 1, "Yes", "No")) %>% 
  ggplot(aes(x = back_problems, fill = back_problems)) +
  geom_bar(aes(y = (..count..) / tapply(..count.., ..PANEL.., sum)[..PANEL..])) +
  scale_y_continuous(labels = scales::percent) +
  facet_wrap(~ sex) +
  labs(
    x = "Back Problems",
    y = "",
    title = "Reported Back Pain",
    subtitle = "Data from students at California Polytechnic State University",
    fill = "Back Pain"
  )
```
```{r}
# alternative solution

backpack %>%
  mutate(
    back_problems = ifelse(
      back_problems == 1, "Yes", "No"
      )
  ) %>%
  ggplot(
    aes(x = sex, fill = back_problems))+
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::percent)+
  scale_fill_manual(values = c("Yes" =  "cornflowerblue", "No" = "darkorchid")) +
  labs(
    y = "Percentage of Students",
    x = "Report of back pain"
  )
```





The percentage of women reporting back problems is much higher than the percentage of men (over the double).


### Question 8

Now you’ve investigated your data separately, create plots to describe the relationship between backpack weight, body weight, gender and reported backpain.

Try both backpack_weight and body_weight as well as the ratio.

Hint: faceting would be useful here.