---
title: "R Notebook"
output: html_notebook
---


```{r}
library(tidyverse)
library(GGally)
library(ggfortify)
library(fastDummies)
```

```{r}
house <- read_csv("data/housing_prices.csv")
house
```  
Notes:  
- What does one row represent? An aggregate of a district in California  
- Some variables are raw sums, some are stats, like averages  
--> do some feature engineering / variablbe transformation to get more stats that are meaningful about these houses  


```{r}
house %>% 
  select(total_rooms, total_bedrooms) %>% 
  ggpairs()
```
- strongly positively skewed (1, 4)  
- strongly positively correlated (2)  
- normal distribution (3)

We could investigate:  
- (rooms - bedrooms) = non-bedrooms  
- bedrooms / rooms = proportion of bedrooms  


Feature Engineering  
- The data set we start with is the "raw material"  
- We can use those to make more meaning and aid interpretation --> "feature engineering"  



```{r}
house <- house %>% 
  mutate(bedrooms_per_house = total_bedrooms / households,
         rooms_per_house = total_rooms / households,
         people_per_house = population / households,
         bedrooms_per_person = total_bedrooms / population,
         rooms_per_person = total_rooms / population,
         bedrooms_per_room = total_bedrooms / total_rooms,
         non_bedrooms_per_house = rooms_per_house - bedrooms_per_house
         )

glimpse(house)

skimr::skim(house) %>% view()
```

* Extra idea: start using long & lats to make more meaning --> joining to external city coordinate data, creating labels, flags, or things like distance from city or from the coast.  


Let's explore the relationships between the variables - what looks like it influences price?


```{r message=FALSE}
ggpairs1 <- house %>% 
  select(median_house_value,
         housing_median_age,
         median_income,
         ocean_proximity) %>% 
  ggpairs()

ggpairs1
```

- house age and house value = weakly positively correlated  
- income and house value = strongly positively correlated
- ocean proximity and house value = 3 categories behave similarly --> what do we do with these? group them?  
- plot 1 shows the distribution of house values across the various districts = skewed


```{r}
house %>% 
  count(ocean_proximity)
```

```{r message=FALSE}
ggpairs2 <- house %>% 
  select(median_house_value,
         total_rooms,
         rooms_per_house,
         total_bedrooms,
         bedrooms_per_house) %>% 
  ggpairs()

ggpairs2
```

- all weakly positively correlated to median_house_value  

```{r message=FALSE}
ggpairs3 <- house %>% 
  select(median_house_value,
         households,
         population,
         people_per_house,
         total_rooms,
         rooms_per_house) %>% 
  ggpairs()

ggpairs3
```

- households --> proxy --> urban/city

```{r}
house %>% 
  select_if(is.numeric) %>% 
  skimr::skim()
```

Might want to do some transformation of highly positively skewed numeric --> e.g. take logs  

```{r}
mod1 <- lm(median_house_value ~ median_income,
           data = house)

summary(mod1)


```

- stars = significance
- is the estimate equals to zero or significantly different? --> price = (constant = a = intercept = 45457) + (increment = b = 39987) x income
- r^2 = explains 41.3% of variation (it goeas from 0 to 1)

```{r}
autoplot(mod1)
```

- 1) residual = how wrong we were, how far from our model --> fitted_value = our estimate of the price = what we guessed; residual = the amount we were wrong by; is hovering around zero in a random way  
- 2) where the points would be if they were normal --> we want our residual to be normal;
doesn't look like a straight line, indication of long tails  
- 3) is the amount we are wrong by consistent? is there  pattern? 2 standard deviations wrong --> a pattern of how wrong we were = yes --> where the points are more clustered we are more wrong (?)  



```{r}
mod2 <- lm(median_house_value ~ ocean_proximity,
           data = house)

summary(mod2)
```

```{r}
mod3 <- lm(median_house_value ~ ocean_proximity + median_income,
           data = house)

summary(mod3)
```

```{r}
library(raster)
library(leaflet)

raster <- rasterFromXYZ(house %>% 
  dplyr::select(longitude, latitude, median_house_value), crs = "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")

pal <- colorNumeric(c("blue", "red", "yellow"), values(raster),
  na.color = "transparent")

leaflet() %>% 
  addTiles() %>% 
  addRasterImage(raster, colors = pal, opacity = 0.7) %>% 
  addLegend(pal = pal, values = values(raster),
    title = "Median House Value")
```

Log transformations  


```{r}
ggplot(house, aes(x = median_income)) +
  geom_histogram()

ggplot(house, aes(x = log(median_income))) +
  geom_histogram()
```


lognormal centered around the mean

```{r}
mod4 <- lm(median_house_value ~ ocean_proximity + 
             log(median_income) + median_income,
           data = house)

summary(mod4)
```

```{r}
ggplot(house, aes(x = median_house_value)) +
  geom_histogram()

ggplot(house, aes(x = log(median_house_value))) +
  geom_histogram()
```

```{r}
mod5 <- lm(log(median_house_value) ~ ocean_proximity + 
             log(median_income) + median_income,
           data = house)

summary(mod5)

plot(mod5)
```

