---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(GGally)
library(ggfortify)
library(fastDummies)
```

```{r}
housing_prices <- read_csv("data/housing_prices.csv")
housing_prices
```

```{r}
# checking if there are NA's
summary(housing_prices)

# imputing NA's with the total_bedrooms mean
housing_prices <- housing_prices %>%
  mutate(total_bedrooms = coalesce(total_bedrooms, median(total_bedrooms, na.rm = TRUE)))

housing_prices
```



```{r message=FALSE}
ggpairs(housing_prices)
```

```{r}
housing_prices <- housing_prices %>% 
  select(-total_bedrooms)

housing_prices
```

r = 0.643 between median_house_value and median_income and possibly ocean_proximity  

- The response is median_house_value  
- The predictor is median_income  

```{r}
housing_prices %>%
  ggplot(aes(x = median_income, y = median_house_value)) +
  geom_point() +
  scale_y_continuous(labels = scales::label_comma()) +
  geom_smooth(method = "lm", se = FALSE) +
  labs()
```

```{r}
lm1 <- lm(median_house_value ~ median_income, data = housing_prices)

autoplot(lm1)

summary(lm1)
```


```{r}
housing_prices %>%
  ggplot(aes(y = median_house_value)) + 
  geom_boxplot() +
  scale_y_continuous(labels = scales::label_comma())
```

The median is around 175000, while the estimated median_house_values are accurate to only 74870 on average (residual standard error)  

```{r}
housing_prices %>% 
  distinct(ocean_proximity)
# 5 distinct values ---> 4 dummy variables
```

```{r}
housing_prices_dummy <- housing_prices %>%
  dummy_cols(select_columns = "ocean_proximity", remove_first_dummy = TRUE, remove_selected_columns = TRUE) %>% 
  janitor::clean_names()

head(housing_prices_dummy)
# the reference category is <1H OCEAN
```

```{r}
housing_prices %>%
  ggplot(aes(x = ocean_proximity, y = median_house_value)) + 
  geom_boxplot() +
  scale_y_continuous(labels = scales::label_comma())
```

There is evidence to investigate whether or not ocean_proximity has an effect on house_value, and as such we think there is an association between the two variables.  

```{r}
housing_prices_dummy %>%
  summarise(correlation = cor(ocean_proximity_island, median_house_value))

housing_prices_dummy %>%
  summarise(correlation = cor(ocean_proximity_near_bay, median_house_value))

housing_prices_dummy %>%
  summarise(correlation = cor(ocean_proximity_near_ocean, median_house_value))

housing_prices_dummy %>%
  summarise(correlation = cor(ocean_proximity_inland, median_house_value))
```
Found one strong negative correlation between ocean_proximity_inland and median_house_value  
